# Data Appendix

This appendix presents descriptive statistics and graphical analyses that support the main empirical results but are not included in the core regression analysis.

---
title: "Education, Institutional Quality, and Environmental Quality in Africa"
author: "EMMANUEL KODOM"
date: "`r Sys.Date()`"
output: word_document
---


```{r}
# Load necessary libraries
#load packages
library(readr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(stargazer)
library(writexl)

# Load the pollution data
pollution <- read_csv("data/clean_pollution_data.csv")
unique_countries <- unique(pollution$country)
unique_countries
length(unique_countries)  # Count of unique countries
colSums(is.na(pollution))

```

Variable: co2_pc

Definition:
Per capita carbon dioxide emissions measured in metric tons per person per year. This is the main dependent variable representing environmental pollution.

Source:
World Bank, World Development Indicators (WDI).

```{r}
#Summary statistics for CO2 emissions per capita
summary(pollution$co2_pc)
hist(pollution$co2_pc, 
     main = "Histogram of CO2 Emissions per Capita", 
     xlab = "CO2 (metric tons per person)", 
     col = "lightblue", border = "white")

```
```{r}
# A boxplot for CO2 emissions per capita
boxplot(pollution$co2_pc,
        main = "Boxplot of CO2 Emissions",
        ylab = "CO2 per capita")

#Time series plot for CO2 emissions per capita
pollution %>%
  group_by(year) %>%
  summarise(avg_co2 = mean(co2_pc, na.rm = TRUE)) %>%
  ggplot(aes(year, avg_co2)) +
  geom_line() +
  labs(title = "Average CO2 Emissions in Africa (2002–2023)",
       y = "CO2 per capita")

```


Variable: edu

Definition:
Human Capital Index (education) measuring the average quality and years of schooling adjusted by returns to education.

Source:
Penn World Table (Version 10.0), Feenstra et al. (2015).

```{r}
# Summary statistics for Education variable
summary(pollution$edu)
hist(pollution$edu, 
     main = "Histogram of Human Capital Index", 
     xlab = "Human Capital (Index)", 
     col = "lightgreen", border = "white")

```




Variable: gov

Definition:
Institutional quality index computed as the average of six Worldwide Governance Indicators (Voice & Accountability, Political Stability, Government Effectiveness, Regulatory Quality, Rule of Law, and Control of Corruption).

Source:
World Bank, Worldwide Governance Indicators (WGI).


```{r}
# Summary statistics for Institutional Quality variable
summary(pollution$gov)
hist(pollution$gov, 
     main = "Histogram of Institutional Quality", 
     xlab = "Institutional Quality Index", 
     col = "orange", border = "white")

```



Variable: gdp_pc

Definition:
Real GDP per capita (constant 2015 U.S. dollars), capturing the level of economic development.

Source:
World Bank, World Development Indicators (WDI).

```{r}
# Summary statistics for GDP per capita variable
summary(pollution$gdp_pc)
hist(pollution$gdp_pc, 
     main = "Histogram of GDP per Capita", 
     xlab = "GDP per Capita (2015 USD)", 
     col = "lightgray", border = "white")

```



Variable: energy_use

Definition:
Per capita energy use measured in kilograms of oil equivalent per person.

Source:
World Bank, World Development Indicators (WDI).


```{r}
# Summary statistics for Energy Use variable
summary(pollution$energy_use)
hist(pollution$energy_use, 
     main = "Histogram of Energy Use per Capita", 
     xlab = "Energy Use (kg of oil equivalent per person)", 
     col = "lightcoral", border = "white")

```



Variable: trade_gdp, industry_gdp, agri_land, urban_share

These variables serve as control variables capturing trade exposure, industrial structure, 
agricultural activity, and urbanization, respectively.

Each is expressed as a percentage of GDP (for trade and industry) or percentage 
of total land/population. All are derived from the World Development Indicators.





```{r}
# Map of African Countries Included in the Study
library(sf)
library(rnaturalearth)
library(ggplot2)
library(dplyr)

# countries
countries <- c(
  "Angola","Benin","Botswana","Burkina Faso","Burundi","Cameroon",
  "Central African Republic","Democratic Republic of the Congo","Republic of the Congo",
  "Cote d'Ivoire","Egypt","Eswatini","Ethiopia","Gabon","Gambia",
  "Ghana","Kenya","Lesotho","Liberia","Madagascar","Malawi","Mali","Mauritius",
  "Morocco","Mozambique","Namibia","Niger","Nigeria","Rwanda","Senegal",
  "Sierra Leone","South Africa","Sudan","Tanzania","Togo","Tunisia","Uganda",
  "Zambia","Zimbabwe"
)

# Load Africa map
africa <- ne_countries(continent = "africa", returnclass = "sf")

# Mark selected countries
africa$selected <- ifelse(africa$name %in% countries, "Included", "Not Included")

# Plot
ggplot(africa) +
  geom_sf(aes(fill = selected), color = "black") +
  scale_fill_manual(values = c("Included" = "orange", "Not Included" = "lightgray")) +
  labs(title = "African Countries Included in the Study",
       fill = "Dataset Status") +
  theme_minimal()



```

```{r}
# ===============================
# CO₂ Emissions Gradient Map for Africa
# ===============================
# ---------------------------------
# 1. Compute average CO2 per capita
# ---------------------------------
co2_avg <- pollution %>%
  group_by(country) %>%
  summarise(avg_co2 = mean(co2_pc, na.rm = TRUE))

# ---------------------------------
# 2. Standardize country names
# ---------------------------------
rename_map <- c(
  "Congo, Dem. Rep."      = "Democratic Republic of the Congo",
  "Congo, Rep."           = "Republic of the Congo",
  "Cote d'Ivoire"         = "Côte d'Ivoire",
  "Gambia, The"           = "Gambia",
  "Egypt, Arab Rep."      = "Egypt",
  "Eswatini"              = "eSwatini"
)

co2_avg$country_clean <- dplyr::recode(co2_avg$country, !!!rename_map)

# ---------------------------------
# 3. Load Africa map
# ---------------------------------
africa <- ne_countries(continent = "africa", returnclass = "sf")

# ---------------------------------
# 4. Merge map with CO2 data
# ---------------------------------
africa_map <- africa %>%
  left_join(co2_avg, by = c("name" = "country_clean"))

# ---------------------------------
# 5. Plot the CO2 gradient map
# ---------------------------------
ggplot(africa_map) +
  geom_sf(aes(fill = avg_co2), color = "black", size = 0.2) +
  scale_fill_viridis_c(option = "inferno", na.value = "gray90",
                       name = "Avg CO2 (tons per capita)") +
  labs(
    title = "Average CO2 Emissions in Africa (2002–2023)",
    subtitle = "Countries color-coded by mean per-capita emissions",
    caption = "Source: WDI (2023)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )

```

